How to select predictive models for decision-making or causal inference?

### 2024-10-20 response_to_reviewer.md

> Eq. (3) (now 4) is an expression of the estimator of the ATE based on a model of the outcome, sometimes called g-estimation. It is unbiased under consistency (Assumption. 3) and unconfoundedness (Assumption 1). To prove that, take the expectation of Eq. (3).  Then apply  Eq. (1) (now 2) which is true under consistency (Eq. 1 and Assumption 1) and unconfoundedness (Assumption 3). Then by the law of total expectancy, the conditionnality on X disappear and we recover Eq. (2) (now 3)

> G-estimation itself does not control error on both populations. The formulas in Eq. (1-3) (now 2-4) are all partly oracle formulas: they rely on conditional expectations but give no specific procedures on how to compute them. We added a clarification on this topic above Eq. (4) (now 5) and above Eq. (8) (now 9).

*Furthermore, the Robinson decomposition (R-decomposition) should be better discussed and its role in the context of causal inference should be better commented.
R-decomposition introduces two quantities, the conditional mean outcome and the probability to be treated (known as propensity score). What is the role of the propensity score? How is it calculated? Is it calculated simply by a frequentist approach (number of treated/total number of patients)?*

> G-computation relies on one choice of decomposition of the CATE
> estimation (Eq. (2)), which leads to into a sequence of regressions,
> where the only statistical estimate is the response function µa(x)
> [Chernozhukov et al., 2024, chapter 15](https://causalml-book.org/). We
> have clarified that this decomposition is just a mathematical
> rewriting. The mean outcome from the treatment effect and the
> propensity score are two statistical quantity that enable this
> rewritting . Any statistical estimator can be chosen to model these
> quantities, in particular in the section were we introduce them, they
> are written in term of expectations, not detailing an estimation
> procedure, as this is discussed later. We have clarfied this below
> formula 4 (formely 3) and in the begining of the section containing
> formula 9 (formely 8). A frequentist approach counting rations of
> patients is one such estimator, but it will work only for a small
> number of discrete covariate. In practice, researchers often estimate
> the propensity score with a logistic regression, but other methods such
> as Boosted trees, Random Forests, or Neural Networks can be used. A
> good ps estimator should be well calibrated but this question is out of
> scope for this paper. In our discussion, we briefly mentioned this
> interesting research question on what properties should have a good
> estimator for the propensity score.

*Finally, I note that the paper mentions inference methods specific to machine learning (random forest, etc.). In this regard, there is a great deal of study in the literature as to their actual usefulness in this field. I suggest the authors consider/review some of the following literature in the paper (those that the authors consider most appropriate to be included in their manuscript):*

@@

https://www.research-collection.ethz.ch/handle/20.500.11850/650703

https://aisel.aisnet.org/pacis2022/181/

or some other that can be easily found in literature.

> We thank the reviewer for the references. We have added two references for Causal ML into the manuscript: [Chernozhukov et al., 2024](https://causalml-book.org/) and [Kaddour et al., 2022](https://arxiv.org/pdf/2206.15475). Both of them cover most of the subfields at the intersection of causal inference and machine learning. I have added them at the beginning of the manuscript before stating our focus on model selection for CATE.

### 2024-10-20 response_to_reviewer.md

> Eq. (3) (now 4) is an expression of the estimator of the ATE based on a model of the outcome, sometimes called g-estimation. It is unbiased under consistency (Assumption. 3) and unconfoundedness (Assumption 1). To prove that, take the expectation of Eq. (3).  Then apply  Eq. (1) (now 2) which is true under consistency (Eq. 1 and Assumption 1) and unconfoundedness (Assumption 3). Then by the law of total expectancy, the conditionnality on X disappear and we recover Eq. (2) (now 3)

> G-estimation itself does not control error on both populations. The formulas in Eq. (1-3) (now 2-4) are all partly oracle formulas: they rely on conditional expectations but give no specific procedures on how to compute them. We added a clarification on this topic above Eq. (4) (now 5) and above Eq. (8) (now 9).

*Furthermore, the Robinson decomposition (R-decomposition) should be better discussed and its role in the context of causal inference should be better commented.
R-decomposition introduces two quantities, the conditional mean outcome and the probability to be treated (known as propensity score). What is the role of the propensity score? How is it calculated? Is it calculated simply by a frequentist approach (number of treated/total number of patients)?*

> G-computation relies on one choice of decomposition of the CATE
> estimation (Eq. (2)), which leads to into a sequence of regressions,
> where the only statistical estimate is the response function µa(x)
> [Chernozhukov et al., 2024, chapter 15](https://causalml-book.org/). We
> have clarified that this decomposition is just a mathematical
> rewriting. The mean outcome from the treatment effect and the
> propensity score are two statistical quantity that enable this
> rewritting . Any statistical estimator can be chosen to model these
> quantities, in particular in the section were we introduce them, they
> are written in term of expectations, not detailing an estimation
> procedure, as this is discussed later. We have clarified this below
> formula 4 (formely 3) and in the begining of the section containing
> formula 9 (formely 8). A frequentist approach counting ratios of
> patients is one such estimator, but it will work only for a small
> number of discrete covariate. In practice, researchers often estimate
> the propensity score with a logistic regression, but other methods such
> as Boosted trees, Random Forests, or Neural Networks can be used. A
> good ps estimator should be well calibrated but this question is out of
> scope for this paper. In our discussion, we briefly mentioned this
> interesting research question on what properties should have a good
> estimator for the propensity score.

@@

*Finally, I note that the paper mentions inference methods specific to machine learning (random forest, etc.). In this regard, there is a great deal of study in the literature as to their actual usefulness in this field. I suggest the authors consider/review some of the following literature in the paper (those that the authors consider most appropriate to be included in their manuscript):*

*https://link.springer.com/book/10.1007/978-3-031-35051-1*

*https://www.causal-machine-learning.com/*

*https://www.frontiersin.org/journals/bioinformatics/articles/10.3389/fbinf.2021.746712/full*

*https://www.research-collection.ethz.ch/handle/20.500.11850/650703*

*https://aisel.aisnet.org/pacis2022/181/*

*or some other that can be easily found in literature.*

> We thank the reviewer for the references. We have added several
> references for Causal ML into the manuscript, in particular [Kaddour et al., 2022](https://arxiv.org/pdf/2206.15475), corresponding to
> https://www.causal-machine-learning.com/, and [Chernozhukov et al., 2024](https://causalml-book.org/) which both cover most of the
> subfields at the intersection of causal inference and machine learning.

### 2024-10-20 how_to_select_latex.tex
Finish address rev 1

\paragraph{Robinson decomposition}
G-computation is a choice of decomposition of the CATE estimation. Other choices
of decomposition exist, such as the R-decomposition
\cite{robinson_rootnconsistent_1988}. The latter introduces two new statistical
estimates, the conditional mean outcome and the probability of being treated (known
as propensity score \cite{rosenbaum_central_1983}):
\begin{flalign}
    \text{Conditional mean outcome} &                    & m(x) \myeq \; & \mathbb E_{Y \sim
    \mathcal{D}} [Y|X=x]            &                    &
    \label{def:m}
    \\
    \text{Propensity score}         &                    &
    e(x) \myeq \;                   & \mathbb P[A=1|X=x]
    \label{def:propensity_score}
\end{flalign}
with these, the outcome (Eq.\,\ref{eq:consistency}) can be written
\begin{flalign}\label{eq:r_decomposition}
    \text{R-decomposition}            &  & y(a) = m(x) + \big( a - e(x) \big)
    \tau(x) + \varepsilon(x; a)\notag &  &
    \\\ &&\text{with}\quad \mathbb E[\varepsilon(X; A)|X, A] = 0&&
\end{flalign}
$m$ and $e$ are often called
\emph{nuisances} \cite{chernozhukov_double_2018}. They are unknown and are estimated in practice thanks to statistical models (eg. linear models or forest based estimators).

>>>>

\paragraph{Robinson decomposition}
G-computation is a choice of decomposition of the CATE estimation. Other choices
of decomposition exist, such as the R-decomposition
\cite{robinson_rootnconsistent_1988}. The latter introduces two new statistical
estimates, the conditional mean outcome and the probability of being treated (known
as propensity score \cite{rosenbaum_central_1983}):
\begin{flalign}
    \text{Conditional mean outcome} &                    & m(x) \myeq \; & \mathbb E_{Y \sim
    \mathcal{D}} [Y|X=x]            &                    &
    \label{def:m}
    \\
    \text{Propensity score}         &                    &
    e(x) \myeq \;                   & \mathbb P[A=1|X=x]
    \label{def:propensity_score}
\end{flalign}
with these, the outcome (Eq.\,\ref{eq:consistency}) can be written
\begin{flalign}\label{eq:r_decomposition}
    \text{R-decomposition}            &  & y(a) = m(x) + \big( a - e(x) \big)
    \tau(x) + \varepsilon(x; a)\notag &  &
    \\\ &&\text{with}\quad \mathbb E[\varepsilon(X; A)|X, A] = 0&&
\end{flalign}
$m$ and $e$ are often called
\emph{nuisances} \cite{chernozhukov_double_2018}. They are unknown and must be estimated from the data.

\bigskip

Both the in the ATE and CATE formula Eq.\,(\ref{eq:ate_estimate},
\ref{eq:cate_estimate}), and the Robinson decomposition involve
conditional expectations --the response functions $\mu_a(x)$ or the
nuisances $m(x)$ and $e(x)$. In practice those are given by statistical
models: linear models, random forests, etc
\cite{kaddour2022causal,chernozhukov2024applied}.
